Choosing TTC Subway Service Provides the Most Time Stability during Weekday Rush-Hours

### 2024-09-24 outputs/paper/paper.qmd
added conent to discussion & fixed figure position in rendered product

Where the data comes from - All data used for and throughout the paper are extracted from Open Data Toronto Portal [@OpenDataToronto]. TTC Bus Delay data [@Bus], TTC Streetcar Delay [@Streetcar], TTC Subway Delay [@Subway] and TTC Subway Delay Codes [@Subway] are retrieved to compare the delay magnitudes of different services. Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], dplyr [@dplyr], hms [@hms], ggplot2 [@ggplot2], patchwork [@patchwork], and knitr [@knitr].
\
\
For all delay dataset, only data in 2023 were chosen. There are three
main reasons for this choice:

@@

```{r}
#| label: fig-rush-off
#| fig-cap: Barplot used to visualize the longest average delay time for different TTC services during Rush-Hours and Off-Peak Hours
#| echo: false
#| message: false

@@

morning_rush <- ggplot(morning_avg_delay_rush, aes(x = service_type, y = morning_mean_delay_rush, fill = service_type)) + geom_bar(stat = "identity") + 
  # Add labels for total hours
  geom_text(aes(label = paste0(round(mean_delay_time, 1), " mins")), 
            vjust = 1, color = "white", size = 2, position = position_stack(vjust = 0.6)) +  
  labs(title = "Morning Rush Average Delay", 
       x = "Service Type", y = "Average Delay Time (min)", fill = "Service Type") +  # Label the axes
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 8))

evening_rush <- ggplot(evening_avg_delay_rush, aes(x = service_type, y = evening_mean_delay_rush, fill = service_type)) + geom_bar(stat = "identity") + 
  # Add labels for total hours
  geom_text(aes(label = paste0(round(mean_delay_time, 1), " mins")), 
            vjust = 1, color = "white", size = 2, position = position_stack(vjust = 0.6)) +  
  labs(title = "Evening Rush Average Weekday Delay", 
       x = "Service Type", y = "Average Delay Time (min)", fill = "Service Type") +  # Label the axes
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))

@@

knitr::kable(reason_delay, col.names = c("Bus", "Streetcar", "Subway"), align = c("c", "c","c"))
```

>>>>

Where the data comes from - All data used for and throughout the paper are extracted from Open Data Toronto Portal [@OpenDataToronto]. TTC Bus Delay data [@Bus], TTC Streetcar Delay [@Streetcar], TTC Subway Delay [@Subway] and TTC Subway Delay Codes [@Subway] are retrieved to compare the delay magnitudes of different services. Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], dplyr [@dplyr], hms [@hms], ggplot2 [@ggplot2], patchwork [@patchwork], kableExtra [@kableExtra], and knitr [@knitr].
\
\
For all delay dataset, only data in 2023 were chosen. There are three
main reasons for this choice:

@@

```{r}
#| label: fig-rush
#| fig-cap: Barplot used to visualize the longest average delay time for different TTC services during Rush-Hours and Off-Peak Hours
#| fig-pos: "H"
#| echo: false
#| message: false

@@

morning_rush <- ggplot(morning_avg_delay_rush, aes(x = service_type, y = morning_mean_delay_rush, fill = service_type)) + geom_bar(stat = "identity") + 
  # Add labels for total hours
  geom_text(aes(label = paste0(round(mean_delay_time, 1), " mins")), 
            vjust = 1, color = "white", size = 2, position = position_stack(vjust = 0.6)) +  
  labs(title = "AM-Rush Average Delay", 
       x = "Service Type", y = "Average Delay Time (min)", fill = "Service Type") +  # Label the axes
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 8))

evening_rush <- ggplot(evening_avg_delay_rush, aes(x = service_type, y = evening_mean_delay_rush, fill = service_type)) + geom_bar(stat = "identity") + 
  # Add labels for total hours
  geom_text(aes(label = paste0(round(mean_delay_time, 1), " mins")), 
            vjust = 1, color = "white", size = 2, position = position_stack(vjust = 0.6)) +  
  labs(title = "PM-Rush Average Delay", 
       x = "Service Type", y = "Average Delay Time (min)", fill = "Service Type") +  # Label the axes
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))

@@

knitr::kable(reason_delay, col.names = c("Bus", "Streetcar", "Subway"), align = c("c", "c", "c")) |>
  kableExtra::kable_styling(position = "center")
```
